---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# jalcal: an R package to convert Jalaali and Gregorian calendar dates

<!-- badges: start -->
![CRAN Downloads](https://cranlogs.r-pkg.org/badges/last-month/jalcal)
<!-- badges: end -->

## Overview

The Persian or Jalaali calendar, also called the Solar Hijri or Hijri Shamsi 
calendar, is a 12-month system. 

- It begins its count from the Hijra, which is July 19, 622 AD in the standard 
Gregorian calendar.

- The Persian New Year begins on Nowruz, the spring equinox. This means the start 
of the year is timed with the start of spring, as determined by astronomical 
observations.

- The first six months have 31 days each, the next five have 30 days, and the 
last month has 29 days in common years and 30 days in leap years. Each season 
consists of three consecutive months.

- The exact leap year determination also relies on the vernal equinox date and 
time. The calculations and conversions in this package are based on the work by 
Kazimierz M. Borkowski (1996) who used an analytical model of Earth's motion 
to calculate equinoxes from AD 550 to 3800 and identify leap years based on 
Tehran time.

- The Persian calendar is the official calandar in Iran and Afghanistan and is 
also used in parts of Tajikistan and Kurdistan, particularly among Persian- and 
Kurdish-speaking communities.


- Borkowski, K. M. (1996). The Persian calendar for 3000 years. 
*Earth, Moon, and Planets*, 74, 223–230. [doi:10.1007/BF00055188](https://doi.org/10.1007/BF00055188)

## Installation

To install the development version of `jalcal` from [GitHub](https://github.com/jalilian/jalcal), use the [remotes](https://CRAN.R-project.org/package=remotes) package:
```r
# Install remotes if not already installed
install.packages("remotes")

# Install getsat from GitHub
remotes::install_github("jalilian/jalcal")
```

To install the package from [CRAN](https://CRAN.R-project.org/package=jalcal), use the `install.packages` command:
```r
# install.packages('jalcal')
```

## Example 1: Convert Jalaali dates to Gregorian dates

To convert **Jalaali dates** (e.g., `1403-12-30` and `1404-01-01`) to **Gregorian dates**, use:
```{r example1}
library("jalcal")
dates <- jal2greg(year=c(1403, 1404), month=c(12, 1), day=c(30, 1))
print(dates)
```


## Example 2: Convert Gregorian dates to Jalaali dates

To convert **Gregorian dates** to **Jalaali dates**, use:
```{r example2}
set.seed(14040101)
gdates <- sample(seq.Date(from = as.Date("561-01-01"), 
                                to = as.Date("3177-12-31"), by = "day"), 
                       size = 100)
jdates <- greg2jal(gdates)
cbind(gdates, jdates)
```

